---
title: "Class09"
author: "Hanqing Zhao"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prepare the data
```{r}
fna.data <- "/Users/zhq/BIMM143/WisconsinCancer.csv"
wisc.df <- read.csv(fna.data)
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[3:32])
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
# Create diagnosis vector for later 
diagnosis <-wisc.df$diagnosis
#diagnosis
```
How many observations in the data? 569
```{r}
dim(wisc.data)
```

How many of the observations have a malignant diagnosis?
```{r}
m<-sum(as.numeric(wisc.df$diagnosis=="M"))
m
```


## Perform PCA
```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = T)
# Look at summary of results
summary(wisc.pr)
biplot(wisc.pr)
```
```{r}
# Scatter plot observations by components 1 and 2
plot( wisc.pr$x, col = c(diagnosis) , 
     xlab = "PC1", ylab = "PC2")
```
Generate pca plot for PC1 and PC3
```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[,1:3], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```

# Hiearchical clustering
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
plot(wisc.hclust)
abline(h = 17, col="red", lty=2)
```

wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters, diagnosis)
grps <- cutree(wisc.pr.hclust)


wisc.km <- kmeans(, centers= , nstart= )


# 3D

library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)


# Prediction
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc


plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")








